---
title: "Project Two"
author: "Mary Brown and Jennifer Relihan"
date: "10/21/2021"
output: html_document
---

## Introduction


## Data  
### Read in the data and learn about the dimensions as well as different column names. 
```{r,echo=TRUE, message=FALSE}  
library(magrittr)
library(dplyr)
Data<-read.csv("OnlineNewsPopularity.csv")  
dim(Data)  
names(Data)  
str(Data)
```  

## Subset  
### Subset the data by channels and create a vector of all of the channels we are interested in.  
```{r,echo=TRUE}  
EntertainmentChannel <- Data %>% filter(data_channel_is_entertainment == TRUE) %>% select(-starts_with("data_channel_is_"))  

AllChannels <- Data %>% select(starts_with("data_channel_is_")) %>% names  
```  

### Check for any missing values.  
```{r,echo=TRUE}  
sum(is.na(Data))  
```  

## Summary Statistics  
### Here we produced some summary statistics and graphs of the data. First, we looked at a full summary on each column of our data. Next, we decided to specifically look further into total shares by days of the week. In order to do summaries on days of the week, we needed to gather the days/values by shares and filter by true values. The first summary shows the standard deviation, average, median, and IQR for each day of the week by shares. The second summary shows the same summary statistics for each day of the week by minimum amount of shares. 
```{r,echo=TRUE, message = FALSE}  
library(tidyverse)  
summary(Data) 

Days <- Data %>% pivot_longer(starts_with("weekday_is"), names_to = "day", values_to = "data") %>% filter(data == TRUE) 

Summary1 <- Days %>% group_by(day) %>% summarize(std = sd(shares), average = mean(shares), median = median(shares), IQR = IQR(shares))  
print(Summary1)

Summary2 <- Days %>% group_by(day) %>% summarize(std = sd(kw_min_min), average = mean(kw_min_min), median = median(kw_min_min), IQR = IQR(kw_min_min))  
print(Summary2)  

```  

## Section for Plots 
```{r,echo=TRUE, message=FALSE, warning =FALSE}  
library(ggplot2)  
library(scales)  

Videos <- ggplot(Data, aes(x = num_videos, y = shares)) + geom_bar(stat = "identity", fill = "tan3") + xlim(-1,21) + scale_y_continuous(labels = unit_format(unit = "M", scale = 5e-6)) + ggtitle("Shares by Videos") + labs(y = "Shares", x = "Number of Videos", caption = "This view shows the number of videos filtered from 0 to 20.") + theme(plot.caption = element_text(hjust = 0))  
print(Videos)

Images <- ggplot(Data, aes(x = num_imgs, y = shares)) + geom_bar(stat = "identity", fill = "steelblue") + xlim(-1,21) + scale_y_continuous(labels = unit_format(unit = "M", scale = 5e-6)) + ggtitle("Shares by Images") + labs(y = "Shares", x = "Number of Images", caption = "This view shows the number of images filtered from 0 to 20.") + theme(plot.caption = element_text(hjust =0))
print(Images) 


# Plot 3:
# This changes word content 0 values to NA. We do this because we want to see the articles with words and how many shares they get. The 0 values had a lot more shares but are comprised of videos or photos with no words.
EntertainmentChannel$n_tokens_content[EntertainmentChannel$n_tokens_content == 0] <- NA
# The number of shares based on the number of words in the content to review how this effects shares.
Num_words <- ggplot(EntertainmentChannel, aes(x=n_tokens_content, y=shares))+ geom_bar(stat = "identity", fill="steelblue") + labs(y="Number of Shares", x="Number of Words") + ggtitle("Shares by Content Size") + theme(plot.caption = element_text(hjust =0))
# Here you can review the trend of shares as a function of the size or word count of an article.
Num_words


# Plot 4:
# Here we are exploring how the rate of positive words in an article effect the amount of shares
positivity <- ggplot(EntertainmentChannel, aes(x=global_rate_positive_words, y=shares))+ geom_point(stat = "identity", fill="steelblue") + labs(y="Number of Shares", x="Rate of Positive Words") + ggtitle("Shares by Positivity") + theme(plot.caption = element_text(hjust =0))
positivity

# Plot 5:
# Here we are exploring how the rate of negative words in an article effect the amount of shares
negativity <- ggplot(EntertainmentChannel, aes(x=global_rate_negative_words, y=shares))+ geom_point(stat = "identity", fill="steelblue") + labs(y="Number of Shares", x="Rate of Negative Words") + ggtitle("Shares by Negativity") + theme(plot.caption = element_text(hjust =0))
negativity

# Putting Plots 4 and 5 together to review side by side.Here you can review the shares by the rate of positive or negative content. Another point of review is to look at the rate of positive or negative words based off of the channel type. For instance, entertainment articles have a max rate of 0.15 positive content and a max rate of only 0.075 for negative words. We can see that the site Mashable tends to write more positive content for entertainment.
pos_neg_join <- ggpubr::ggarrange(positivity, negativity,ncol=2)
pos_neg_join
```  

## Modeling  
### In this section we split the data - 70% train and 30% test  
```{r,echo=TRUE, message=FALSE} 
library(caret)  
set.seed(123)
DataIndex<-createDataPartition(y = Data$shares, p = 0.7, list = FALSE)  
TrainData <- Data[DataIndex,]  
TestData <- Data[-DataIndex,]  
```  




