---
title: "Project Two"
author: "Mary Brown and Jennifer Relihan"
date: "10/21/2021"
output: html_document
params:
  channel: "Entertainment"
---

=======

## Introduction


## Required to run this document  
```{r, echo = TRUE}  
library(tidyverse)  
library(caret)  
library(ggplot2)  
library(ggpubr)  
library(magrittr)  
library(scales)
library(corrplot)  
library(GDAtools)
library(gbm)
library(shiny)
```  

## Data Manipulation 
```{r,echo=TRUE, message=TRUE} 
# Here we read in the data and learn more about the dimensions as well as different column names.
Data<-read.csv("OnlineNewsPopularity.csv")  
dim(Data)  
names(Data)  
str(Data)  

# After reviewing the attributes information on the website, we need to remove the non-predictive variables which are URL and Time Delta. Then, we need to subset the data by channels.  
Data <- Data %>% select(-url, -timedelta)  
EntertainmentChannel <- Data %>% filter(data_channel_is_entertainment == TRUE) %>% select(-starts_with("data_channel_is_"))  

# Check for any missing values  
sum(is.na(Data))  

# In order to work with days of the week, we created a column that shows all of the days of the week by shares.  
Days = c('Sunday', 'Monday', 'Tuesday', 'Wednesday', 'Thursday', 'Friday', 'Saturday')  
EntertainmentChannel<- EntertainmentChannel %>% mutate(weekday = ifelse(weekday_is_monday==1, 'Monday',  
                                    ifelse(weekday_is_tuesday==1, 'Tuesday',   
                                    ifelse(weekday_is_wednesday==1, "Wednesday", 
                                    ifelse(weekday_is_thursday==1, "Thursday",
                                    ifelse(weekday_is_friday==1, "Friday",
                                    ifelse(weekday_is_saturday==1, "Saturday", "Sunday"  
                                           ))))))) %>% mutate(weekday = factor(weekday, levels = Days))  

# We would like to create a popularity column by shares for future analysis. This popularity column will rate by either "Popular" or "Unpopular." To do this, we found the median value for the amount of shares. 
ShareSummary <- EntertainmentChannel %>% summarize(median = median(shares))  
EntertainmentChannel <- EntertainmentChannel %>% mutate(Popularity = ifelse(shares>1200, "Popular","Unpopular"))  

# Before doing any exploratory data analysis (EDA), we need to split the data -70% train and 30% test.
NumericData <- dplyr::select_if(EntertainmentChannel, is.numeric)
set.seed(123)
DataIndex<-createDataPartition(y = EntertainmentChannel$shares, p = 0.7, list = FALSE)  
TrainData <- EntertainmentChannel[DataIndex,]
TestData <- EntertainmentChannel[-DataIndex,]  
dim(TrainData)  
dim(TestData)
```  


## Exploratory Data Analysis  
```{r,echo=TRUE, message=FALSE}  
# This table shows summary statistics for our 'shares' data point. 
SharesStats <- TrainData %>% summarize(Avg = mean(shares), Median = median(shares), STDV = sd(shares), IQR = IQR(shares))
knitr::kable(SharesStats, caption = "Summary Statistics on Shares", digits = 0)

# This table shows the summary statistics for shares by weekday. 
Weekdays <- TrainData %>% group_by(weekday) %>% summarise(Avg = round(mean(shares)), Maximum = round(max(shares)), Minimum = round(min(shares))) 
knitr::kable(Weekdays, caption = "Summary Statistics on Shares by Weekdays", digits = 2)  

# Showing popularity for the days of the week. A designation of Popular is when the number of shares is greater than 1200, and less than 1200 is considered Unpopular.
weekday_popularity <- table(TrainData$weekday, TrainData$Popularity)
weekday_popularity

# Showing popularity for weekdays vs weekends.For the y-axis column 0 = weekday and 1 = weekend. Here we can summarize popularity of content based on if it was shared over the weekend or during the week. 
is_weekend_pop <- table(TrainData$is_weekend, TrainData$Popularity)
is_weekend_pop

# This is a table displaying the average number of shares of referenced articles in mashable, by weekday. 
GDAtools::wtable(TrainData$weekday, w = TrainData$self_reference_avg_sharess)

# This is a table displaying average keywords on certain days of the week  
Keywords <- TrainData %>% group_by(weekday) %>% summarise(Average = mean(self_reference_avg_sharess), Median = median(self_reference_avg_sharess), STDV = sd(self_reference_avg_sharess), IQR = IQR(self_reference_avg_sharess))  
knitr::kable(Keywords, caption = "Summary statistics of average shares of referenced articles by specific days of the week", digits = 2)
```  

### Plots 
```{r,echo=TRUE, message=FALSE, warning =FALSE}  
# Plot 1: 
# This plot displays the number of shares by weekday. 
DaysPlot <- TrainData %>% ggplot(aes(x = weekday, y = shares)) + geom_bar(stat = "identity", fill = "steelblue") + scale_y_continuous(labels = unit_format(unit = "M", scale = 5e-6)) + ggtitle("Shares by Days of the Week") + labs(y = "Shares", x = "Days of the Week") + theme(axis.text.x = element_text(angle = 45))
print(DaysPlot)  

# Plot 2:
# This plot displays the popularity of shares by days of the week.
PopularityPlot <- TrainData %>% ggplot(aes(x = weekday)) + geom_bar(aes(fill = as.factor(Popularity))) + labs(x = "Popularity by Day") + theme(axis.text.x = element_text(angle = 45)) + scale_fill_discrete(name = "Popularity")  
print(PopularityPlot)

# Plot 3:
# This plot displays shares by the number of Videos.
Videos <- TrainData %>% ggplot(aes(x = num_videos, y = shares)) + geom_bar(stat = "identity", fill = "tan3") + xlim(-1,21) + scale_y_continuous(labels = unit_format(unit = "M", scale = 5e-6)) + ggtitle("Shares by Videos") + labs(y = "Shares", x = "Number of Videos", caption = "This view shows the number of videos filtered from 0 to 20.") + theme(plot.caption = element_text(hjust = 0))  
print(Videos)  

# Plot 4:
# This plot displays shares by the number of images. 
Images <- TrainData %>% ggplot(aes(x = num_imgs, y = shares)) + geom_bar(stat = "identity", fill = "steelblue") + xlim(-1,21) + scale_y_continuous(labels = unit_format(unit = "M", scale = 5e-6)) + ggtitle("Shares by Images") + labs(y = "Shares", x = "Number of Images", caption = "This view shows the number of images filtered from 0 to 20.") + theme(plot.caption = element_text(hjust =0))
print(Images) 

# Plot 5:
# This changes word content 0 values to NA. We do this because we want to see the articles with words and how many shares they get. The 0 values had a lot more shares but are comprised of videos or photos with no words.
TrainData$n_tokens_content[TrainData$n_tokens_content == 0] <- NA
# The number of shares based on the number of words in the content to review how this effects shares.
Num_words <- ggplot(TrainData, aes(x=n_tokens_content, y=shares))+ geom_bar(stat = "identity", fill="steelblue") + labs(y="Number of Shares", x="Number of Words") + ggtitle("Shares by Content Size") + theme(plot.caption = element_text(hjust =0))
# Here you can review the trend of shares as a function of the size or word count of an article.
Num_words

# Plot 6:
# Here we are exploring how the rate of positive words in an article effect the amount of shares
positivity <- ggplot(TrainData, aes(x=global_rate_positive_words, y=shares))+ geom_point(stat = "identity", fill="steelblue") + labs(y="Number of Shares", x="Rate of Positive Words") + ggtitle("Shares by Positivity") + theme(plot.caption = element_text(hjust =0))
positivity

# Plot 7:
# Here we are exploring how the rate of negative words in an article effect the amount of shares
negativity <- ggplot(TrainData, aes(x=global_rate_negative_words, y=shares))+ geom_point(stat = "identity", fill="steelblue") + labs(y="Number of Shares", x="Rate of Negative Words") + ggtitle("Shares by Negativity") + theme(plot.caption = element_text(hjust =0))
negativity

# Putting Plots 4 and 5 together to review side by side.Here you can review the shares by the rate of positive or negative content. Another point of review is to look at the rate of positive or negative words based off of the channel type. For instance, entertainment articles have a max rate of 0.10 positive content and a max rate of 0.093 for negative words in this training data set. We can see that the site Mashable tends to write more positive content for entertainment.
pos_neg_join <- ggpubr::ggarrange(positivity, negativity,ncol=2)
pos_neg_join
```  

## Linear Regression Models  
### Supervised learning includes regression models, tree based methods, and k nearest neighbors. The basic linear regression model includes a response, a value of our explanatory variable for the ith observation, the y-intercept, and the slope. The model aims to show a linear approach for modeling the relationship between predictors and some response. The model is fit by minimizing the sum of squared residuals, which is equivalent to assuming normality on  errors and using maximum liklihood to estimate the beta's. In R, the basic linear model fits done with lm(). When utilizing lm() in R, statistical analysis can be found using anova() or summary(). 
```{r,echo=TRUE}  
# A summary on the full model shows predictors with significant p-values. I selected predictors with significant p-values to further explore for my linear regression model.  
FullData <- lm(shares~., data = TrainData)  
summary(FullData) 

# This is the model I chose with some significant predictors 
Model1<-as.formula("shares ~ n_unique_tokens + kw_max_min + kw_avg_min + n_non_stop_unique_tokens + num_self_hrefs + num_keywords + kw_min_max + kw_max_avg + global_subjectivity + abs_title_sentiment_polarity + n_tokens_title + global_sentiment_polarity")  

# Fitting the model with the training data 
fit1 <- train(Model1, data = TrainData, 
              method = "lm",  
              preProcess = c("center", "scale"),  
              trControl = trainControl(method = "cv", number = 10))  

# Making a table of the results from the first linear model 
Results <- data.frame(t(fit1$results))  
colnames(Results) <- "Model Results"
knitr::kable(Results, digits = 3)  

# View Results
Results

# Linear Regression Model 2:

# After talking with a subject matter expert on marketing we discussed the columns available in the data set and which ones they may be most interested in seeing in a model to predict shares. 

# Fitting the model with the training data.  
fit2 <- train(shares ~weekday + global_rate_positive_words + global_rate_negative_words + title_sentiment_polarity,  
              data = TrainData,  
              method = "lm",   
              preProcess = c("center", "scale"),  
              trControl = trainControl(method = "cv", number = 10))  

# Putting together the results of Model 2
results <- data.frame(t(fit2$results))  
colnames(results) <- "Model Results"  
knitr::kable(results, digits = 3)  

# View results  
results  

# Putting both Linear Regression models together  
lmr_results <- data.frame(t(fit1$results), t(fit2$results))  
names(lmr_results)[1] <- "Model 1 Results"  
names(lmr_results)[2] <- "Model 2 Results"  
knitr::kable(lmr_results, digits = 3)  

# View results  
lmr_results
```  

## Random Forest Model  
### Random Forest modeling falls under supervised learning. It builds an ensemble of decision trees which help to get a more accurate prediction. The RF model extends the idea of bagging, generally better than bagging. It creates multiple trees from bootstrap samples. 
```{r, echo = TRUE, cache = TRUE, warning = FALSE}  
# Converting NA values back to 0
TrainData[is.na(TrainData)] = 0  

# Fitting the model using repeated cross validation
rfFit <- train(shares ~.,   
               data = select(TrainData, -Popularity),
               method = "rf",  
               preProcess = c("center", "scale"),  
               trControl = trainControl(method = "repeatedcv", number = 5, repeats = 3),  
               tuneGrid = data.frame(mtry = seq(1,10,1)))  

# View resampling results across tuning parameters 
rfFit  
```  


## Boosted Tree Model
### The boosted tree model is a way to slowly train a tree so that you don't overfit. Here we use cross validation to determine the best tuning parameters - shrinkage(λ), n.trees(B), and interaction.depth (related to d). Then we take those tuning parameters and fit the model.Then we predict shares on the test data using the best model. 
```{r}

# Parameter tuning using 10-fold cross validation to determine parameter values for fitting the boosted tree model
fitControl <- trainControl(
                          method = "repeatedcv",
                          number = 10,
                          repeats = 10)

# Fitting the model using repeated cross-validation
boostFit <- train(shares ~ .,
                  data= select(TrainData, -Popularity),
                  method="gbm",
                  trControl = fitControl,
                  verbose = FALSE)
# View re-sampling results across tuning parameters
boostFit
# Take the best tuning results and save as an object
bestTune <- boostFit$bestTune
n.tress_best <- bestTune$n.trees
interaction.depth_best <- bestTune$interaction.depth
shrinkage_best <- bestTune$shrinkage

# Re-fit the model using the best tuning results
boostFit_best <- gbm(shares ~ .,
                     data=select(TrainData, -Popularity),
                     distribution = "gaussian",
                     n.trees = n.tress_best,
                     shrinkage = shrinkage_best,
                     interaction.depth = interaction.depth_best)
```  

## Predictions 
### Predictions with the 4 models 
```{r, echo=TRUE, cache = TRUE, message = FALSE}  
# Prediction on the first linear model  
pred <- predict(fit1, newdata = TestData)  
A <- postResample(pred, obs = TestData$shares)

pred2 <- predict(fit2, newdata = TestData)  
B <- postResample(pred2, obs = TestData$shares)  

pred3 <- predict(rfFit, newdata = select(TestData, -shares))  
C <- postResample(pred3, obs = TestData$shares)

pred4 <- predict(boostFit_best, newdata = select(TestData, -shares), n.trees = n.tress_best)  
D <- postResample(pred4, obs = TestData$shares)  

# A table of all prediction results  
Predictions <- t(rbind(A[1], B[1], C[1], D[1]))  
colnames(Predictions) <- c("Linear Model 1", "Linear Model 2", "Random Forest Model", "Boosted Tree Model")  
Predictions  
```  

## Comparison
### RMSE comparison to declare a model winner  
```{r, echo=TRUE}  
# RMSE comparison
LR1results <- sqrt(mean((pred - TestData$shares)^2))
LR2results <- sqrt(mean((pred2 - TestData$shares)^2))  
RFresults <- sqrt(mean((pred3 - TestData$shares)^2))
BTresults <- sqrt(mean((pred4 - TestData$shares)^2))  

# Make a table of results from the above  
FinalResults <- rbind(LR1results, LR2results, RFresults, BTresults)
row.names(FinalResults2) <- c("Linear 1", "Linear 2", "RF Model", "BT Model")  
knitr::kable(FinalResults2, digits = 4, caption = "This table shows RMSE comparisons of each model. The model with the lowest RMSE value is declared the winning model.")
```  




